On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2018-02-02 manuscript.Rmd
minor edits; adjusted figures; eamus reference

@@

| Variable      |  Definition                                                                                       |
| ------------- | ------------------------------------------------------------------------------------------------- |
| g~cut~        | Conductance of an isolated non-stomatal cuticle.                                                  |
| g~min~        | Minimum conductance measured from weight loss of detached leaves.                                 |
| g~dark~       | Night-time conductance, or conductance after significant dark adaptation.                         |
| g~0~          | Intercept in the Ball-Berry-type stomatal conductance model, i.e. g~s~ when A~n~ approaches zero. |
| g~s,low A~    | Cuvette measurements of g~s~ when photosynthesis rates are very low.                              |

\
Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1). The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996]. We further discuss cuticular conductance in Section V.

@@

\
We quantitatively compared each of the different 'minimum' conductance (Table 1) values by synthesizing existing data. To estimate g~0~ (Eq. 1) from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata). Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see SuppInfo 2 for references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Supporting Information for full details of the compilation). All conductance values were converted to per unit projected surface area. Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance (discussed further in Section VI), it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the @lin2015 database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the 52 studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted or extrapolated g~0~ term.

\
The results of the quantitative comparison are presented in Fig. 1. From this compilation, a few striking differences between estimates are apparent. Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^). This result is in line with the methods comparison of @kerstiens1996. More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cut~. However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis. Nonetheless, the comparison between g~cut~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.

@@

\
**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but a few data points occur outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cut~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq. 1) with regression; g~s,low A~ leaf conductance during periods of very low photosynthesis. See text for data sources and methods.

@@

\
In the Ball-Berry type framework of stomatal conductance models, the effects of photosynthetic photon flux density (PPFD) and leaf temperature on g~s~ are both assumed to enter via the dependence of A~n~ on these drivers. Here we show that the g~0~ parameter not only sets a minimum value of g~s~ in the model (when A~n~ is zero), it also modifies the behaviour of Eq. 1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model'). @leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995]. Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model. Despite these early reports, the exact role of g~0~ in models is often overlooked.

\
In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015]. This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2016]. @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata. Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cut~ (Fig. 1), we ignore this effect in the model simulations that follow.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry model, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata. Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018, @urban2017].

@@

\
**Figure 4**. Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq. 1) on leaf fluxes. (a) Intrinsic water-use efficiency (A~n~/g~s~) as a function of he photosynthetic photon flux density (PPFD), holding other environmental drivers constant, for three values of g~0~. (b) The same simulations as in (a), but showing the intercellular CO~2~ concentration (C~i~). (c) Stomatal conductance (g~s~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. (d) The same simulations as in (c), but showing C~i~. Note how C~i~ increases at high VPD and T~air~, only when g~0~> 0. For all simulations it is assumed that T~leaf~ is equal to T~air~, and ignoring the differential permeability of the cuticle to CO~2~ and H~2~O [@hanson2016].

@@

\
Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle. Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996]. The main barrier to diffusion is actually a very thin layer at the leaf surface. Because most of the resistance is located in such a thin layer, g~cut~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b]. Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (**cite**), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010]. If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?

@@

\
Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged. Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but was able to explain only some of the wide range of g~cut~ measured across different plant species. Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.

@@

\
A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition when net photosynthesis is zero. However as our literature comparison showed (Fig. 1), g~dark~ is substantially higher compared to any other measure of minimum conductance (except fitted g~0~). The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.

@@

\
Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories. Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ or g~cut~ (pooled) by plant growth form. One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), and g~min~. Such correlation did not bear out across all species in our database, and interesting to note is that *Eucalyptus* species (n = 11, included in *Myrtales*) - all measured on the driest continent of Australia - have g~min~ values slightly higher than the average (Fig. 6b).

@@

\
While there is clearly plenty of variation in g~min~ between species, a number of lines of evidence suggest that g~min~ also has great potential to vary within species. Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age and with altitude. This discussion is directly relevant to the use of g~min~ in models, because if intra-specific variation is large, it complicates model parameterization.

@@

\
A number of studies have attempted to attribute the drought-induced acclimation in g~min~ to a change in the chemical composition of the cuticle. @bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and an increase in the amount of cuticular waxes, but could not find a relationship between the two and the response was highly genotype-specific. However, @premachandra1992 found that epicuticular wax load increased on leaves of non-irrigated Sorghum cultivars and was positively correlated with cuticular conductance and cell membrane stability. @mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts). @bi2017 also reported that drought caused a change in cuticular wax production and chemical composition, but again in a highly genotype-specific way. Thus it can be concluded that the cuticle indeed changes substantially after a change in plant available water, but in a complex, species-specific manner that is yet to be connected directly to changes g~min~. This opens up the possibility that leaky stomata are a significant component of the drought response, not just the cuticular composition.

@@

\
The response of g~min~ to temperature is more complex; it shows both a response to instantaneous changes in temperature, as well as an acclimatory response to growth temperature. @riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cut~ to temperature, with the response becoming especially steep at higher temperatures (> 40C). The mechanism of the temperature response is complex, highly species-specific, and we refer to @schuster2016b for a detailed investigation. The quick increase in g~min~ at high temperature may well be a crucial component of the ability of plants to tolerate heat waves [@drake2018]. In support of the link between g~min~ and heat tolerance, @schuster2016a reported on a negative between thermal tolerance and g~min~ across nine species, so that species with improved tolerance to very high temperature had a low g~min~.

\
Less well established is the acclimatory potential of g~min~ to changes in growth temperature. We present a case study on *Eucalyptus parramattensis* grown in whole-tree chambers (see SuppInfo 3 for methods details). The chambers either tracked ambient conditions, or were subjected to a persistent +3C warming treatment. After months of growth in the treatments, g~min~ was measured at various instantaneous temperatures, ranging from 17.5 to 27.5C. We found a 56% decrease in g~min~ in the elevated temperature trees, but there was no clear pattern with measurement temperature (Fig. 8). This decrease in g~min~ is consistent with the drought response, because again g~min~ is reduced in leaves that are subjected to conditions that increase evaporative demand (vapour pressure deficit was higher in the elevated temperature treatment, see @drake2018). The direction of this response is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time. Responses across temperature gradients are more complex, in particular the change in g~min~ with altitude has been well studied. It is commonly  reported that g~min~ increases with altitude [e.g. @delucia1984; @herrick1991; @anfodillo2002]. @fernandez2017 discuss the literature on altitude responses in detail and argue that the short growing seasons at high altitude are insufficient for complete maturation of leaf cuticles.

@@

$$ g_s = g_{min} + (g_0 - g_{min}) + g_1 \frac{A_n}{C_a}f(D)$$

>>>>

- bel looks for data on low PAR; maybe replace gslowA with gslowPAR, so we can argue that g0 is much higher than gslowPAR. 
- some data behind figure 4a (and maybe Arnaud 4d?)
- min(g0, gs) ('Uddling', has he used it?)  vs. g0 + gs.

We could say that plant dilemma is to be both impermeable to water loss (because they are saturated with water and leave are in an extremely dry air environment and hence need a external layer to prevent dessication= cuticle) AND need to be permeable to CO2 to grow. The dilemma is that leaves cannot be permeable to CO2 and not to H20. Hence the evolution of stomata that disrupt the cuticle layer enabling CO2 entre but cannot avoid some CO2 exit but can also close to limit H20. loss.

  It could be interested to discuss somewhere in the text the relevance of the conductance (gmin) versus the flow (Emin). Obviously gmin is more an intrinsic property of the leaf but what matters for the leaf water relation of the flow. In terms of drought resistance and water balance (and water relations) Ecuti is more relevant than gmin I think. At night for instance gdark is high but vpd so the flow might not be so different than at gmin during the day.

finding of Schuster e al (2016) who found a high proportion of triterpenoid in the cuticule of a desertic thermotolerant species. This compound would act as to stabilize the cuticule.

@@

| Variable      |  Definition                                                                                       |
| ------------- | ------------------------------------------------------------------------------------------------- |
| g~cuti~        | Conductance of an isolated non-stomatal cuticle.                                                  |
| g~min~        | Minimum conductance measured from weight loss of detached leaves.                                 |
| g~dark~       | Night-time conductance, or conductance after significant dark adaptation.                         |
| g~0~          | Intercept in the Ball-Berry-type stomatal conductance model, i.e. g~s~ when A~n~ approaches zero. |
| g~s,low A~    | Cuvette measurements of g~s~ when photosynthesis rates are very low.                              |

\
Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1). The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cuti~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996]. We further discuss cuticular conductance in Section V.

@@

\
We quantitatively compared each of the different 'minimum' conductance (Table 1) values by synthesizing existing data. To estimate g~0~ (Eq. 1) from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata). Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see SuppInfo 2 for references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Supporting Information for full details of the compilation). All conductance values were converted to per unit projected surface area. Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance (discussed further in Section VI), it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the @lin2015 database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the 52 studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted or extrapolated g~0~ term.

\
The results of the quantitative comparison are presented in Fig. 1. From this compilation, a few striking differences between estimates are apparent. Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^). This result is in line with the methods comparison of @kerstiens1996. More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cuti~. However in the same study, a literature compilation of hundreds of values showed that g~cuti~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cuti~ were pooled in their analysis. Nonetheless, the comparison between g~cuti~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cuti~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.

@@

\
**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but a few data points occur outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq. 1) with regression; g~s,low A~ leaf conductance during periods of very low photosynthesis. See text for data sources and methods.

@@

\
In the Ball-Berry type framework of stomatal conductance models, the effects of photosynthetic photon flux density (PPFD) and leaf temperature on g~s~ are both assumed to enter via the dependence of A~n~ on these drivers. Here we show that the g~0~ parameter not only sets a minimum value of g~s~ in the model (when A~n~ is zero), it also modifies the behaviour of Eq. 1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model'). @leuning1990 defined g~0~ as g~s~ when A~n~ approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A~n~/g~s~ (and C~i~) at low light [see also @leuning1995]. Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model. Despite these early reports, the exact role of g~0~ in models is often overlooked.

\
In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015]. This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2016]. @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata. Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cuti~ (Fig. 1), we ignore this effect in the model simulations that follow.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry model, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cuti~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata. Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018, @urban2017].

@@

\
**Figure 4**. Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq. 1) on leaf fluxes. (a) Intrinsic water-use efficiency (A~n~/g~s~) as a function of he photosynthetic photon flux density (PPFD), holding other environmental drivers constant, for three values of g~0~. (b) The same simulations as in (a), but showing the intercellular CO~2~ concentration (C~i~). (c) Leaf transpiration (E~L~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. (d) The same simulations as in (c), but showing C~i~. Note how C~i~ increases at high VPD and T~air~, only when g~0~> 0. For all simulations it is assumed that T~leaf~ is equal to T~air~, and ignoring the differential permeability of the cuticle to CO~2~ and H~2~O [@hanson2016].

@@

\
Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle. Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996]. The main barrier to diffusion is actually a very thin layer at the leaf surface. Because most of the resistance is located in such a thin layer, g~cuti~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b]. Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (**cite**), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010]. If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?

@@

\
Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged. Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but was able to explain only some of the wide range of g~cuti~ measured across different plant species. Similarly, @hauke1998 found no relationship between the gradual decrease in g~cuti~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.

@@

\
A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition when net photosynthesis is zero. However as our literature comparison showed (Fig. 1), g~dark~ is substantially higher compared to all other variables except fitted g~0~. The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.

@@

\
Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories. Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ or g~cuti~ (pooled) by plant growth form. One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), and g~min~. Such correlation did not bear out across all species in our database, and interesting to note is that *Eucalyptus* species (n = 11, included in *Myrtales*) - all measured on the driest continent of Australia - have g~min~ values slightly higher than the average (Fig. 6b).

@@

\
While there is clearly plenty of variation in g~min~ between species, a number of lines of evidence suggest that g~min~ also has great potential for plasticity. Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age and with altitude. This discussion is directly relevant to the use of g~min~ in models, because if the degree of plasticity is large, it complicates model parameterization.

@@

\
A number of studies have attempted to attribute the drought-induced acclimation in g~min~ to a change in the chemical composition of the cuticle. @bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and an increase in the amount of cuticular waxes, but could not find a relationship between the two and the response was highly genotype-specific. However, @premachandra1992 found that epicuticular wax load increased on leaves of non-irrigated Sorghum cultivars and was positively correlated with cuticular conductance and cell membrane stability. @mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts). @bi2017 also reported that drought caused a change in cuticular wax production and chemical composition, but again in a highly genotype-specific way. Thus it can be concluded that the cuticle indeed changes substantially after a change in plant available water, but in a complex, species-specific manner that is yet to be connected directly to changes g~min~. This opens up the possibility that leaky stomata are a significant component of the drought response, not just the composition of the cuticle.

@@

\
The response of g~min~ to temperature is more complex; it shows both a response to instantaneous changes in temperature, as well as an acclimatory response to growth temperature. @riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cuti~ to temperature, with the response becoming especially steep at higher temperatures (> 40C). In *Eucalyptus haemastoma*, the response of g~min~ to temperature was so steep, that the proportion of cuticular to total transpiration increased from 2-3% at 20C, to 40% at 38C [@eamus2008]. The mechanism of the instantaneous temperature response is complex, highly species-specific, and we refer to @schuster2016b for a detailed investigation. The quick increase in g~min~ at high temperature may well be a crucial component of the ability of plants to tolerate heat waves [@drake2018]. In support of the link between g~min~ and heat tolerance, @schuster2016a reported on a negative between thermal tolerance and g~min~ across nine species, so that species with improved tolerance to very high temperature had a low g~min~.

\
Less well established is the acclimatory potential of g~min~ to changes in growth temperature. We present a case study on *Eucalyptus parramattensis* grown in whole-tree chambers (see SuppInfo 3 for methods details). The chambers either tracked ambient conditions, or were subjected to a persistent +3C warming treatment. After several months of growth in the treatments, g~min~ was measured at various temperatures, ranging from 17.5 to 27.5C. We found a 56% decrease in g~min~ in the elevated temperature trees, but there was no clear pattern with measurement temperature (Fig. 8). This decrease in g~min~ is consistent with the drought response, because again g~min~ is reduced in leaves that are subjected to conditions that increase evaporative demand (vapour pressure deficit was higher in the elevated temperature treatment, see @drake2018). The direction of this response is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time. Responses across temperature gradients are more complex, in particular the change in g~min~ with altitude has been well studied. It is commonly  reported that g~min~ increases with altitude [e.g. @delucia1984; @herrick1991; @anfodillo2002]. @fernandez2017 discuss the literature on altitude responses in detail and argue that the short growing seasons at high altitude are insufficient for complete maturation of leaf cuticles.

@@

$$ g_s = min(g_{min}, g_0) + g_1 \frac{A_n}{C_a}f(D)$$
